Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices

### 2022-07-22 formalization.tex
Fixed quote marks in code

Having described how we adapt the equational theory to model \mEdhoc,
we now move on to the modelling of the adversary model and the 
environment in
which the protocol is executed.
%
We extend the built-in Dolev-Yao adversary rules which are part of 
\mTamarin.
%
We use the following rules to capture the link between a party's identity and
their long-term key pairs, in the \mSig{}- and in the \mStat{}-based methods
respectively.
\begin{center}
\begin{minipage}{0.48\textwidth}
\begin{small}
\begin{verbatim}
1 rule registerLTK_SIG:
2    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
3        [!LTK_SIG($A, ~ltk),
4         !PK_SIG($A, pk(~ltk)),
5         Out(<$A, pk(~ltk)>)]
\end{verbatim}
\end{small}
\end{minipage}
\hfill\vline\hfill
\begin{minipage}{0.48\textwidth}
\begin{small}
\begin{verbatim}
1 rule registerLTK_STAT:
2    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
3        [!LTK_STAT($A, ~ltk),
4         !PK_STAT($A, `g'^~ltk),
5         Out(<$A, `g'^~ltk>)]
\end{verbatim}
\end{small}
\end{minipage}
\end{center}

@@

\begin{minipage}{\textwidth}
\begin{scriptsize}
\begin{verbatim}
1 rule R2_STAT_SIG:
2 let
3    agreed = <CS0, CI, ~CR>
4    gx = `g'^xx
5    data_2 = <`g'^~yy, CI, ~CR>
6    m1 = <`STAT', `SIG', CS0, CI, gx>
7    TH_2 = h(<$H0, m1, data_2>)
8    prk_2e = extr(`e', gx^~yy)
9    prk_3e2m = prk_2e
10   K_2m = expa(<$cAEAD0, TH_2, `K_2m'>,
11               prk_3e2m)
12   protected2 = $V // ID_CRED_V
13   CRED_V = pkV
14   extAad2 = <TH_2, CRED_V>
15   assocData2 = <protected2, extAad2>
16   MAC_2 = aead(`e', K_2m, assocData2, $cAEAD0)
17   authV = sign(<assocData2, MAC_2>, ~ltk)
18   plainText2 = <$V, authV>
19   K_2e = expa(<$cAEAD0, TH_2, `K_2e'>, prk_2e)
20   K_2e_1 = expa(<$cAEAD0, TH_2, `K_2e', `1'>, prk_2e)
21   K_2e_2 = expa(<$cAEAD0, TH_2, `K_2e', `2'>, prk_2e)
22   CIPHERTEXT_2 = <$V XOR K_2e_1, authV XOR K_2e_2>
23   m2 = <data_2, CIPHERTEXT_2>
24   exp_sk = <gx^~yy>
25 in
26   [!LTK_SIG($V, ~ltk), !PK_SIG($V, pkV), In(m1), Fr(~CR), Fr(~yy), Fr(~tid)]
27   --[ExpRunningR(~tid, $V, exp_sk, agreed), R2(~tid, $V, m1, m2)]->
28   [StR2_STAT_SIG($V, ~ltk, ~yy, prk_3e2m, TH_2, CIPHERTEXT_2, gx^~yy, ~tid, m1, m2, agreed),
29    Out(m2)]
\end{verbatim}
\end{scriptsize}
\end{minipage}
\vspace{5mm}

@@

\begin{small}
\begin{verbatim}
1 lemma secrecyPFS:
2    all-traces
3    ``All u v sk tid #t3 #t2.
4        (K(sk)@t3 & CompletedRun(u, v, sk, tid)@t2) ==>
5        ( (Ex #t1. LTKRev(u)@t1 & #t1 < #t2)
6        | (Ex #t1. LTKRev(v)@t1 & #t1 < #t2)
7        | (Ex #t1. EphKeyRev(sk)@t1)
8        | (Ex m1 #t0 #k. I1(tid, u, v, m1)@t0 & TEE(v)@k & (t0 < k | k < t2))
9        | (Ex m1 m2 #t0 #k. R2(tid, v, m1, m2)@t0 & TEE(u)@k & (t0 < k | k < t2))
10       )"
\end{verbatim}
\end{small}
%

>>>>

Having described how we adapt the equational theory to model \mEdhoc,
we now move on to the modelling of the adversary model and the 
environment in
which the protocol is executed.
%
We extend the built-in Dolev-Yao adversary rules which are part of 
\mTamarin.
%
We use the following rules to capture the link between a party's identity and
their long-term key pairs, in the \mSig{}- and in the \mStat{}-based methods
respectively.
\begin{center}
\begin{minipage}{0.48\textwidth}
\begin{scriptsize}
\begin{verbatim}
1 rule registerLTK_SIG:
2    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
3        [!LTK_SIG($A, ~ltk),
4         !PK_SIG($A, pk(~ltk)),
5         Out(<$A, pk(~ltk)>)]
\end{verbatim}
\end{scriptsize}
\end{minipage}
\hfill\vline\hfill
\begin{minipage}{0.48\textwidth}
\begin{scriptsize}
\begin{verbatim}
1 rule registerLTK_STAT:
2    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
3        [!LTK_STAT($A, ~ltk),
4         !PK_STAT($A, `g'^~ltk),
5         Out(<$A, `g'^~ltk>)]
\end{verbatim}
\end{scriptsize}
\end{minipage}
\end{center}

@@

\begin{minipage}{\textwidth}
\begin{scriptsize}
\begin{verbatim}
1 rule R2_STAT_SIG:
2 let
3    agreed = <CS0, CI, ~CR>
4    gx = 'g'^xx
5    data_2 = <'g'^~yy, CI, ~CR>
6    m1 = <'STAT', 'SIG', CS0, CI, gx>
7    TH_2 = h(<$H0, m1, data_2>)
8    prk_2e = extr('e', gx^~yy)
9    prk_3e2m = prk_2e
10   K_2m = expa(<$cAEAD0, TH_2, 'K_2m'>,
11               prk_3e2m)
12   protected2 = $V // ID_CRED_V
13   CRED_V = pkV
14   extAad2 = <TH_2, CRED_V>
15   assocData2 = <protected2, extAad2>
16   MAC_2 = aead('e', K_2m, assocData2,
                  $cAEAD0)
17   authV = sign(<assocData2, MAC_2>, ~ltk)
18   plainText2 = <$V, authV>
\end{verbatim}
\end{scriptsize}
\egroup
\end{minipage}
\begin{minipage}{0.02\textwidth}
\end{minipage}
\begin{minipage}{0.48\textwidth}
\bgroup
\parindent 0pt
\begin{scriptsize}
\begin{verbatim}
19   K_2e = expa(<$cAEAD0, TH_2,
                 'K_2e'>, prk_2e)
20   K_2e_1 = expa(<$cAEAD0, TH_2,
                   'K_2e', '1'>, prk_2e)
21   K_2e_2 = expa(<$cAEAD0, TH_2,
                   'K_2e', '2'>, prk_2e)
22   CIPHERTEXT_2 = <$V XOR K_2e_1,
                     authV XOR K_2e_2>
23   m2 = <data_2, CIPHERTEXT_2>
24   exp_sk = <gx^~yy>
25 in
26   [!LTK_SIG($V, ~ltk), !PK_SIG($V, pkV), In(m1), Fr(~CR), Fr(~yy), Fr(~tid)]
27   --[ExpRunningR(~tid, $V, exp_sk, agreed), R2(~tid, $V, m1, m2)]->
28   [StR2_STAT_SIG($V, ~ltk, ~yy, prk_3e2m, TH_2, CIPHERTEXT_2, gx^~yy, ~tid, m1, m2, agreed),
29    Out(m2)]
\end{verbatim}
\end{scriptsize}
\end{minipage}
\vspace{5mm}

@@

\begin{small}
\begin{verbatim}
1 lemma secrecyPFS:
2    all-traces
3    ''All u v sk tid #t3 #t2.
4        (K(sk)@t3 & CompletedRun(u, v, sk, tid)@t2) ==>
5        ( (Ex #t1. LTKRev(u)@t1 & #t1 < #t2)
6        | (Ex #t1. LTKRev(v)@t1 & #t1 < #t2)
7        | (Ex #t1. EphKeyRev(sk)@t1)
8        | (Ex m1 #t0 #k. I1(tid, u, v, m1)@t0 & TEE(v)@k & (t0 < k | k < t2))
9        | (Ex m1 m2 #t0 #k. R2(tid, v, m1, m2)@t0 & TEE(u)@k & (t0 < k | k < t2))
10       )''
\end{verbatim}
\end{small}
%
